---
title: "Shiny_challenge"
format: html
---

## Shiny Challenge 🌐

Can you write an R script to create a shiny app with the functionality shown in the video below utilizing the [county_data.gpkg](https://github.com/NSF-ALL-SPICE-Alliance/DS421-Carto-Design/blob/main/data/county_data.gpkg) (click download raw file on github after following the hyperlink)and the following packages. Please name this shiny_app_language.qmd and push to your github repository
```{r}
library(shiny)
library(mapview)
library(here)
library(sf)
library(tidyverse)
library(leaflet)
```

### Use this example from class to help you from the fire data
```{r}
ui <- fluidPage(
  titlePanel("Fire Predictions"),
  sidebarLayout(
    sidebarPanel(
      selectInput(
        "var",
        "Choose fire confidence",
        choices = unique(sf_fire_csv_data$confidence)
      )
    ),
    mainPanel(
      leafletOutput("map", height = "700px")
    )
  )
)

server <- function(input, output, session) {
  output$map <- renderLeaflet({
    filtered_data <- sf_fire_csv_data[sf_fire_csv_data$confidence == input$var, ]
    
    mapview(
      filtered_data,
      zcol = "confidence",   # column name, fixed
      layer.name = "Confidence"
    )@map
  })
}

shinyApp(ui, server)
```



### Tips

- This is a .gpkg file (new to us) but you can read it in with st_read()
- The dataframe should have everything you need, no need for new calculations/columns

<https://github.com/user-attachments/assets/5f5327e4-7a42-4439-8350-0fc6c0be064f>


#Challenge
```{r}
county_data <- st_read(here("data/county_data.gpkg")) 
```
```{r}
ui <- fluidPage(
  titlePanel("Language Speakers by County"),
  sidebarLayout(
    sidebarPanel(
      selectInput(
        "var",
        "Select A Language",
        choices = unique(county_data$language)
      )
    ),
    mainPanel(
      leafletOutput("map", height = "700px")
    )
  )
)
 
server <- function(input, output, session) {
  output$map <- renderLeaflet({
    filtered_data <- county_data[county_data$language == input$var, ]
 
    mapview(
      filtered_data,
      zcol = "speakers",       
      layer.name = "# of Speakers"
    )@map
  })
}
 
shinyApp(ui, server)
```

